Specific leaf mass explains difference in biomass partitioning between woody plant functional types

### 2014-10-13 manuscript/draftpaper.Rnw
Made Rnw to replace .tex, added table (for test)
Added variance partioning tables and described in text

\author{Remko Duursma, Daniel Falster and BAAD data contributors}

@@

@@

\item Variance partitioning with R2 for linear mixed-effects models as reported by XXX.

@@

Note: stats done as well, but not yet incorporated into the manuscript.

@@

\item We do note that there was substantial variation between individual plants (Figs.~\ref{fig:figureSI3} and~\ref{fig:figureSI3}). Nonetheless, PFT explained most variation in MF/Asba (to be added). If we also included SLM in the model, PFT was no longer significant. PFT explained very little variation when we partitioned the variance in AF/Asba.

\item We found similar results for the whole-plant leaf mass fraction (LMF), where a clear and strong separation was found between plant functional types (Fig.~\ref{fig:figure3}). See also Fig.~\ref{fig:figureSI5} for the raw data. For the leaf area ratio, pfts were generally closer together except for small evergreen gymnosperms (< 1m height) for unknown reasons. Variance partitioning also confirmed that the differences between PFT could be largely explained by their differences in SLM (not shown yet!).

\item We found no effect of biome on the ratio MF/Asba or AF/Asba (tropical, temperate, boreal) (Fig.~\ref{fig:figureSI2}).

@@

\item We found no effects of climate on anything. This is different from \cite{reich_temperature_2014} but they studied whole-stand biomass, not individual plant biomass. So perhaps this means the climate effects are on stand density not individual plant effects. No effect of climate is good news for models that parameterize the globe by PFT, as this seems sufficient, based on our dataset.

@@

\section{Figures}

>>>>

\author{Remko Duursma, Daniel Falster and BAAD contributors}

@@

<<preload, echo=FALSE>>=
load("tables/Tables_varpart.RData")
@

@@

\item Variance partitioning with marginal (fixed effects) and conditional (fixed and random effects) R2 for linear mixed-effects models as reported by \cite{nakagawa_general_2013}.

@@

@@

\item We do note that there was substantial variation between individual plants (Figs.~\ref{fig:figureSI3} and~\ref{fig:figureSI3}). We estimated the variance explained by PFT, plant height, and biome in MF/Asba using a linear mixed-effects model (see Table~\ref{varpartpipe}). We found that plant height and biome explained only \Sexpr{100*signif(Table_pipemodel_varpart$Marginal[1],2)}\%, but when PFT was added to the model, the variance explained increased to \Sexpr{100*signif(Table_pipemodel_varpart$Marginal[2],2)}\%. If we used specific leaf mass instead of PFT, the R2 was similar (\Sexpr{100*signif(Table_pipemodel_varpart$Marginal[3],2)}\%), showing that PFT affected MF/Asba largely through differences in SLM. Consistent with this, PFT, plant height and biome explained very little variance in ALF/Astba (Table~\ref{varpartpipe}).

\item We found similar results for the whole-plant leaf mass fraction (LMF), where a clear and strong separation was found between plant functional types (Fig.~\ref{fig:figure3}). See also Fig.~\ref{fig:figureSI5} for the raw data. For the leaf area ratio, pfts were generally closer together except for small evergreen gymnosperms (< 1m height) for unknown reasons.

\item Results for variance partitioning of LMF and LAR gave comparable results (Table~\ref{varpartpipe}): PFT increased the variance explained in LMF from \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[1],2)} to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[2],2)}\%. However, SLM was not a good substitute for PFT, as the explained variance decreased to \Sexpr{100*signif(Table_LMFLAR_varpart$Marginal[3],2)}\%. The leaf area ratio, similarly to ALF/Astba, did not depend strongly on PFT or SLA.

\item Although biome was a significant predictor in all variables tested, it consistently explained very little variation (Table~\ref{varpartpipe} and~\ref{varpartlmflar}). Pair-wise comparisons between biomes within PFT never showed significant differences for MF/Asba or AL/Asba (for example boreal vs. temperate evergreen gymnosperm) (Fig.~\ref{fig:figureSI2}).

@@

\item We found no effects of climate on anything. This is different from \cite{reich_temperature_2014} but they studied whole-stand biomass, not individual plant biomass. So perhaps this means the climate effects are on stand density not individual plant effects. No effect of climate is good news for models that parameterize the globe by PFT, as this seems sufficient, based on our dataset. Are there experimental data that show temperature effects on allocation? What about water availability?

@@

\section{Tables}

<<echo=FALSE, message=FALSE, warning=FALSE>>=
library(xtable)
@

<<results='asis', echo=FALSE>>=
print(xtable(Table_pipemodel_varpart, label="varpartpipe", 
             caption="Explained variance in the ratio of leaf mass to basal stem area, and leaf area to basal stem area. The marginal R2 is that explained by the fixed effects only (H = plant height, B = biome, PFT = plant functional type. The conditional R2 is that explained by fixed and random effects (species within study). All fixed effects were highly significant (P \\textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models."), 
      include.rownames=FALSE, caption.placement="top")
@

<<results='asis', echo=FALSE>>=
load("tables/Tables_varpart.RData")
print(xtable(Table_LMFLAR_varpart, label="varpartlmflar",
      caption="Explained variance in the leaf mass fraction (leaf mass / aboveground biomass) and leaf area ratio (leaf area / aboveground biomass). See Table~\\ref{varpartpipe} for more explanation. For plant height, we also added the squared term as a fixed effect, and the interaction with PFT or SLM, because the relationship was non-linear."),
      include.rownames=FALSE, caption.placement="top")
@

\clearpage
\section{Figures}

### 2014-10-13 manuscript/draftpaper.tex
Added variance partioning tables and described in text

\documentclass[a4paper]{article}

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}

\begin{document}

@@

\item We do note that there was substantial variation between individual plants (Figs.~\ref{fig:figureSI3} and~\ref{fig:figureSI3}). Nonetheless, PFT explained most variation in MF/Asba (to be added). If we also included SLM in the model, PFT was no longer significant. PFT explained very little variation when we partitioned the variance in AF/Asba.

\item We found similar results for the whole-plant leaf mass fraction (LMF), where a clear and strong separation was found between plant functional types (Fig.~\ref{fig:figure3}). See also Fig.~\ref{fig:figureSI5} for the raw data. For the leaf area ratio, pfts were generally closer together except for small evergreen gymnosperms (< 1m height) for unknown reasons. Variance partitioning also confirmed that the differences between PFT could be largely explained by their differences in SLM (not shown yet!).

\item We found no effect of biome on the ratio MF/Asba or AF/Asba (tropical, temperate, boreal) (Fig.~\ref{fig:figureSI2}).

@@

\section{Figures}

>>>>

\documentclass[a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\item We do note that there was substantial variation between individual plants (Figs.~\ref{fig:figureSI3} and~\ref{fig:figureSI3}). We estimated the variance explained by PFT, plant height, and biome in MF/Asba using a linear mixed-effects model (see Table~\ref{varpartpipe}). We found that plant height and biome explained only 0.85\%, but when PFT was added to the model, the variance explained increased to 2.6\%. If we used specific leaf mass instead of PFT, the R2 was similar (29\%), showing that PFT affected MF/Asba largely through differences in SLM. Consistent with this, PFT, plant height and biome explained very little variance in ALF/Astba (Table~\ref{varpartpipe}).

\item We found similar results for the whole-plant leaf mass fraction (LMF), where a clear and strong separation was found between plant functional types (Fig.~\ref{fig:figure3}). See also Fig.~\ref{fig:figureSI5} for the raw data. For the leaf area ratio, pfts were generally closer together except for small evergreen gymnosperms (< 1m height) for unknown reasons.

\item Results for variance partitioning of LMF and LAR gave comparable results (Table~\ref{varpartpipe}): PFT increased the variance explained in LMF from 65 to 66\%. However, SLM was not a good substitute for PFT, as the explained variance decreased to 75\%. The leaf area ratio, similarly to ALF/Astba, did not depend strongly on PFT or SLA.

\item Although biome was a significant predictor in all variables tested, it consistently explained very little variation (Table~\ref{varpartpipe} and~\ref{varpartlmflar}). Pair-wise comparisons between biomes within PFT never showed significant differences for MF/Asba or AL/Asba (for example boreal vs. temperate evergreen gymnosperm) (Fig.~\ref{fig:figureSI2}).

@@

\section{Tables}

% latex table generated in R 3.1.1 by xtable 1.7-4 package
% Mon Oct 13 15:43:27 2014
\begin{table}[ht]
\centering
\caption{Explained variance in the ratio of leaf mass to basal stem area, and leaf area to basal stem area. The marginal R2 is that explained by the fixed effects only (H = plant height, B = biome, PFT = plant functional type. The conditional R2 is that explained by fixed and random effects (species within study). All fixed effects were highly significant (P \textless 0.001). In addition to the fixed effects shown, all interactions were added to each of the models.} 
\label{varpartpipe}
\begin{tabular}{llrr}
  \hline
Variable & Predictors & Marginal & Conditional \\ 
  \hline
LMF\_ASTBA & H & 0.01 & 0.53 \\ 
   & H, B & 0.03 & 0.55 \\ 
   & PFT, H, B & 0.29 & 0.64 \\ 
   & SLM, H, B & 0.31 & 0.66 \\ 
  LAR\_ASTBA & H & 0.01 & 0.52 \\ 
   & H, B & 0.05 & 0.53 \\ 
   & PFT, H, B & 0.06 & 0.54 \\ 
   & SLM, H, B & 0.12 & 0.56 \\ 
   \hline
\end{tabular}
\end{table}

% latex table generated in R 3.1.1 by xtable 1.7-4 package
% Mon Oct 13 15:43:27 2014
\begin{table}[ht]
\centering
\caption{Explained variance in the leaf mass fraction (leaf mass / aboveground biomass) and leaf area ratio (leaf area / aboveground biomass). See Table~\ref{varpartpipe} for more explanation. For plant height, we also added the squared term as a fixed effect, and the interaction with PFT or SLM, because the relationship was non-linear.} 
\label{varpartlmflar}
\begin{tabular}{llrr}
  \hline
Variable & Predictors & Marginal & Conditional \\ 
  \hline
LMF & H & 0.65 & 0.89 \\ 
   & H, B & 0.66 & 0.89 \\ 
   & PFT, H, B & 0.75 & 0.88 \\ 
   & SLM, H, B & 0.68 & 0.89 \\ 
  LAR & H & 0.69 & 0.89 \\ 
   & H, B & 0.73 & 0.90 \\ 
   & PFT, H, B & 0.76 & 0.90 \\ 
   & SLM, H, B & 0.77 & 0.92 \\ 
   \hline
\end{tabular}
\end{table}

\clearpage
\section{Figures}
